---
title: "A dicision tree applied to Titanic data set"
author: "Andrey Ziyatdinov"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: united
    toc: true
    keep_md: true
---

```{r options, echo = F}
opts_chunk$set(fig.path = "figures/", comment = NA, results = 'markup', tidy = F, message = F, warning = F, echo = T, cache = T)
```

```{r inc, cache = FALSE, echo = FALSE}
library(plyr)
library(ggplot2)
library(gridExtra)

library(pander)

library(data.table)
```

```{r settings, cache = FALSE, echo = FALSE}
theme_set(theme_light())
panderOptions('table.split.table', Inf)
panderOptions('knitr.auto.asis', FALSE)
```

# Include

Library that implements the decision tree algorithm:

```{r, cache = F}
library(rpart)
```

# Parallel

```{r parallel, echo = T}
library(doMC)
registerDoMC(cores = 2)
```

# Data

```{r dat, echo = T}
test <- fread("~/git/variani/kaggle/titanic/data/test.csv", data.table = F)
train <- fread("~/git/variani/kaggle/titanic/data/train.csv", data.table = F)
```

## Cleaning

```{r cleaning, echo = T}
train <- within(train, {
  Embarked[Embarked == ""] <- NA
})
```

## Transform

```{r dat2, echo = T}
mutate2_dat <- function(df) 
{
  within(df, { 
    Pclass <- ordered(Pclass)  
    Sex <- factor(Sex)
    SibSp <- ordered(SibSp)
    Parch <- ordered(Parch, levels = c("0", "1", "2", "3", "4", "5", "6", "9"))
    Embarked <- factor(Embarked)
  })
}

train2 <- mutate2_dat(train)
test2 <- mutate2_dat(test) 
```


# Model 1: recursive partitioning tree (rpart)

```{r}
fit1 <- rpart(Survived ~ Pclass + Sex + Age + SibSp + Parch + Fare + Embarked, 
  data = train, method = "class")
```

Plot the tree:

```{r plot_fit1}
plot(fit1)
text(fit1)
```

Performance on the training set:

```{r perf1}
tab <- table(train$Survived, predict(fit1, train, type = "class"))
tab

# accuracy
sum(diag(tab)) / sum(tab)
```

Performance on the trainig set was shown [here](http://trevorstephens.com/post/72923766261/titanic-getting-started-with-r-part-3-decision)
to be `0.78`.

Make a submission by yourself.

```{r sub1}
pred1 <- data.frame(PassengerId = test$PassengerId, Survived = predict(fit1, test, type = "class"))
write.csv(pred1, file = "pred1.csv", row.names = FALSE)
```

Result: `0.78469` (#2030, as of 20 April, 2016)

# Model 2: recursive partitioning tree (rpart) + prunning

```{r fit2}
Yt <- as.factor(train$Survived)
Xt <- subset(train, select = c("Pclass", "Sex", "Age", "SibSp", "Parch", "Fare", "Embarked"))

trControl <- trainControl(method = "repeatedcv", number = 5, repeats = 10, selectionFunction = "tolerance") 

set.seed(1)
fit2 <- train(Xt, Yt, trControl = trControl, method = "rpart", tuneLength = 10)

mod2 <- fit2$finalModel
```

```{r fit2_show}
fit2
plot(fit2)
```

When applying `predict` function to prunned tree `fit2`, the best model is used.

```{r check}
table(predict(fit2, test), predict(fit2$finalModel, test, type = "class"))
```

Make a submission #2.

```{r sub2}
pred2 <- data.frame(PassengerId = test$PassengerId, Survived = predict(fit2, test))
write.csv(pred2, file = "pred2.csv", row.names = FALSE)
```

Result: `0.78469` (exactly the same as #1).

```{r check_pred12}
table(pred1$Survived, pred2$Survived)
```

Plotting the model #2 shows that it is rather simple.

```{r plot_mod2}
plot(mod2)
text(mod2)
```

# Model 3: recursive partitioning tree (ctree) + prunning


```{r fit3}
Yt <- as.factor(train2$Survived) 
Xt <- subset(train2, select = c("Pclass", "Sex", "Age", "SibSp", "Parch", "Fare", "Embarked"))

trControl <- trainControl(method = "repeatedcv", number = 5, repeats = 10, selectionFunction = "best") 

set.seed(1)
fit3 <- train(Xt, Yt, trControl = trControl, method = "ctree", tuneLength = 10)

mod3 <- fit3$finalModel
```

```{r plot_mod3, fig.width = 12, fig.height = 9}
plot(mod3) 
```

Make a submission #3.

```{r sub3}
Xp <- subset(test2, select = c("Pclass", "Sex", "Age", "SibSp", "Parch", "Fare", "Embarked"))
pred3 <- data.frame(PassengerId = test$PassengerId, Survived = predict(fit3, Xp))
write.csv(pred3, file = "pred3.csv", row.names = FALSE)
```

```
0.79426

3	Fri, 20 May 2016 16:11:23

Your Best Entry
You improved on your best score by 0.00957. 
You just moved up 648 positions on the leaderboard.
```


```{r check_pred13}
table(pred1$Survived, pred3$Survived)
```

